---
title: "Assignments"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```



This page will contain all the assignments you submit for the class.



### Instructions for all assignments

I want you to submit your assignment as a PDF, so I can keep a record of what the code looked like that day. I also want you to include your answers on your personal GitHub website. This will be good practice for editing your website and it will help you produce something you can keep after the class is over.

1. Download the Assignment1.Rmd file from Canvas. You can use this as a template for writing your answers. It's the same as what you can see on my website in the Assignments tab. Once we're done with this I'll edit the text on the website to include the solutions.

2. On RStudio, open a new R script in RStudio (File > New File > R Script). This is where you can test out your R code. You'll write your R commands and draw plots here.

3. Once you have finalized your code, copy and paste your results into this template (Assignment 1.Rmd). For example, if you produced a plot as the solution to one of the problems, you can copy and paste the R code in R markdown by using the ` ``{r} ``` ` command. Answer the questions in full sentences and Save.

4. Produce a PDF file with your answers. To do this, knit to PDF (use Knit button at the top of RStudio), locate the PDF file in your docs folder (it's in the same folder as the Rproj), and submit that on on Canvas in Assignment 1.

5. Build Website, go to GitHub desktop, commit and push. Now your solutions should be on your website as well.






# Assignment 1

**Collaborators: Carmen Avery, Rachel Villari, and Halle Wasser. **

This assignment is due on Canvas on Monday 9/20 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.


### Problem 1 

Install the datasets package on the console below using `install.packages("datasets")`. Now load the library.

```{r}
datasets::USArrests
install.packages("datasets")
library("datasets")
USArrests
dat<-USArrests # renames USArrests as dat
dat
```
Load the USArrests dataset and rename it `dat`. Note that this dataset comes with R, in the package datasets, so there's no need to load data from your computer. Why is it useful to rename the dataset?

Answer: It's useful to rename the dataset so that we know specifically which dataset we are working with. Additionally, when we edit the dataset it is helpful to have a differentiation (in name) from the original dataset. Additionally "dat" is shorter than the original dataset name.

```{r}

```

### Problem 2

Use this command to make the state names into a new variable called State. 

```{r, eval=FALSE}
dat$state <- tolower(rownames(USArrests))
names(dat)
# This line creates a new variable: states
```

This dataset has the state names as row names, so we just want to make them into a new variable. We also make them all lower case, because that will help us draw a map later - the map function requires the states to be lower case.

List the variables contained in the dataset `USArrests`.
Answer: The variables are murder, assault, urbanpop, and rape. 
```{r}

```

### Problem 3 

What type of variable (from the DVB chapter) is `Murder`? 

Answer: According to the DVB chapter, murder is a quantitative variable. 

What R Type of variable is it?

Answer:In R, Murder is a character variable. 

### Problem 4

What information is contained in this dataset, in general? What do the numbers mean? 

Answer: This dataset lists the average rate of arrests in each state for a specific type of crime. For example, in Iowa there has been an average of 2.2 arrests for murder. The fact that some of the data are listed as decimal numbers indicates that the dataset is reporting averages, as there cannot actually be 2.2 real arrests. 

### Problem 5

Draw a histogram of `Murder` with proper labels and title.

```{r}
hist(dat$Murder, main="US Arrests for Murder", xlab="Average Arrests", ylab="Frequency of Average Arrests") # This line contains the code for the histogram, including its title and axis names
```

### Problem 6

Please summarize `Murder` quantitatively. What are its mean and median? What is the difference between mean and median? What is a quartile, and why do you think R gives you the 1st Qu. and 3rd Qu.?

Answer: Murder's mean is 7.788. Its median is 7.25. The mean is the average of all values in the dataset, whereas the median is simply the middle value of those listed from smallest to largest. The median is also synonymous with the second quartile, Q2. A quartile is when the values of a dataset are divided into four groups. R gives the first and third quartiles because these are common values used in statistics, so it is helpful for R to calculate them automatically. 
```{r}
a<-c(13.2, 10.0, 8.1, 8.8, 9.0, 7.9, 3.3, 5.9, 15.4, 17.4, 5.3, 2.6, 10.4, 7.2, 2.2, 6.0, 9.7, 15.4, 2.1, 11.3, 4.4, 12.1, 2.7, 16.1, 9.0, 6.0, 4.3, 12.2, 2.1, 7.4, 11.4, 11.1, 13.0, 0.8, 7.3, 6.6, 4.9, 6.3, 3.4, 14.4, 3.8, 13.2, 12.7, 3.2, 2.2, 8.5, 4.0, 5.7, 2.6, 6.8) # This line creates the label "a" for the set of data values in Murder. 
mean(a) # This code calculates the mean of a. 
median(a) # This code calculates the median of a. 
```

### Problem 7

Repeat the same steps you followed for `Murder`, for the variables `Assault` and `Rape`. Now plot all three histograms together. You can do this by using the command `par(mfrow=c(3,1))` and then plotting each of the three. 

```{r, echo = TRUE, fig.width = 5, fig.height = 8}
b<-c(236, 263, 294, 190, 276, 204, 110, 238, 335, 211, 46, 120, 249, 113, 56, 115, 109, 249, 83, 300, 149, 255, 72, 259, 178, 109, 102, 252, 57, 159, 285, 254, 337, 45, 120, 151, 159, 106, 174, 279, 86, 188, 201, 120, 48, 156, 145, 81, 53, 161) # creates the label "b" for the set of data values in Assault
mean(b) # calculates the mean of Assault
median(b) # calculates the median of Assault
b<-c(236, 263, 294, 190, 276, 204, 110, 238, 335, 211, 46, 120, 249, 113, 56, 115, 109, 249, 83, 300, 149, 255, 72, 259, 178, 109, 102, 252, 57, 159, 285, 254, 337, 45, 120, 151, 159, 106, 174, 279, 86, 188, 201, 120, 48, 156, 145, 81, 53, 161) # creates the label "c" for the set of data values in Rape
mean(b) # calculates the mean of Rape
median(b) # calculates the median of Rape
hist(dat$Assault, main="US Arrests for Assault", xlab="Average Arrests", ylab="Frequency of Average Arrests") # Code for histogram of Assault, including title and axis labels
hist(dat$Rape, main="US Arrests for Rape", xlab="Average Arrests", ylab="Frequency of Average Arrests") # Code for histogram of Rape, including title and axis labels
par("mar")
par(mar=c(1,1,1,1)) # command to fix the parameters to fix the fact that the figure margins were too large
par(mfrow=c(3,1)) # command to plot multiple histograms
hist(dat$Murder, main="US Arrests for Murder", xlab="Average Arrests", ylab="Frequency of Average Arrests")
hist(dat$Assault, main="US Arrests for Assault", xlab="Average Arrests", ylab="Frequency of Average Arrests")
hist(dat$Rape, main="US Arrests for Rape", xlab="Average Arrests", ylab="Frequency of Average Arrests")
```
Answer: The mean of Assault is 170.76 and the median is 159. The mean of Rape is 21.232 and the median is 20.1. 

What does the command par do, in your own words (you can look this up by asking R `?par`)?

Answer: The command par allows you to set parameters when graphing, or to simply look up certain graphical parameters. 

What can you learn from plotting the histograms together?

Answer:By plotting the histograms together, it becomes easier to see how the rates of arrests compare for the type of crimes committed. For example, it is evident from looking at the histograms that the rate of arrests for assault is much higher than that of murder. It is a helpful way to learn how the difrferent variables interact. 
  
### Problem 8

In the console below (not in text), type `install.packages("maps")` and press Enter, and then type `install.packages("ggplot2")` and press Enter. This will install the packages so you can load the libraries.




# Assignment 2
## Crim 250: Statistics for the Social Sciences
Name: Elizabeth Stoner
Date: 09/27/2021


### Instructions: Copy your code, paste it into a Word document, and turn it into Canvas. You can turn in a .docx or .pdf file. Show any EDA (graphical or non-graphical) you have used to come to this conclusion.

### Problem 1: Load data

Set your working directory to the folder where you downloaded the data.
setwd("C:/Users/Elizabeth/Desktop/CRIM 250")

Read the data
```{r}
#Sys.which("make")
## "C:\\rtools40\\usr\\bin\\make.exe"
#install.packages("jsonlite", type = "source")
#install.packages('plyr', repos = "http://cran.us.r-project.org")
```

```{r}
library(readr)
dat_nsduh_small_1 <- read_csv("dat.nsduh.small.1.csv")
dat <- read_csv("dat.nsduh.small.1.csv")
```



What are the dimensions of the dataset? 
dim(dat)
The dimensions of the dataset are 171 observations of 7 variables. 

names(dat)
The names of the variables are mjage, cigage, iralcage, age2, sexatract, speakengl, and irsex. 


### Problem 2: Variables

Describe the variables in the dataset.
class(dat$mjage)
class(dat$cigage)
class(dat$iralcage)
class(dat$age2)
class(dat$sexatract)
class(dat$speakengl)
class(dat$irsex)
There are seven variables in the dataset. In R, each variable is numeric.   Mjage stands for the age at which respondents used marijuana or hashish for the first time.  Cigage shows the ages at which respondents first started smoking cigarettes every day.  Iralcage is the ages at which respondents first tried alcohol.  Age2 is the most recent age of the respondents. Irsex is the gender of the respondents and is a categorical variable.  Sexatract lists the sexual attraction of the respondents. Speakengl lists how well the respondents speak English. 

What is this dataset about? Who collected the data, what kind of sample is it, and what was the purpose of generating the data?
The dataset is from the National Survey of Drug Use and Health. The dataset is used to understand the extent of drug use and other health issues throughout the United States. The data were collected by the Substance Abuse and Mental Health Services Administration, part of the U.S. Department of Health and Human Services. The NSDUH uses a stratified sample within the United States population. The purpose of gathering this data is to provide an understanding of the health issues pervasive in America in order to establish helpful treatment programs and inform public policy. 


### Problem 3: Age and gender

What is the age distribution of the sample like? Make sure you read the codebook to know what the variable values mean.
```{r}
hist(dat$age2,
     main="Final Age",
     xlab="Age Response Code",
     ylab="Frequency")
```


The majority of the age variables were listed between 12 and 16. According to the codebook, this indicates that most respondents were between 24 and 64 years old. The most frequent ages reported were from 30 to 64 years old. The least frequent ages reported were between 15 to 19 years old. In this dataset there were no ages younger than 15. 

Do you think this age distribution is representative of the US population? Why or why  not?
I would expect to see a higher frequency of 18 to 30 year olds, rather than have the peak start at 30.  Additionally, according to the codebook, the ages of respondents when the questionnaire began were different from when it ended. As a result, I expect that this distribution is not entirely reflective of the age distribution of the current U.S. population; however, it is likely not too skewed either. 

Is the sample balanced in terms of gender? If not, are there more females or males?
```{r}
table(dat$irsex)
```


According to the table generated by R, there are 91 males and 80 females in the dataset. Therefore, it is not balanced as there are more males than females. 

Use this code to draw a stacked bar plot to view the relationship between sex and age. What can you conclude from this plot?
```{r}
tab.agesex <- table(dat$irsex, dat$age2)
barplot(tab.agesex,
        main = "Stacked barchart",
        xlab = "Age category", ylab = "Frequency",
        legend.text = rownames(tab.agesex),
        beside = FALSE) # Stacked bars (default)
```


This plot shows, as would be expected, that there were more men in each age group, aside from ages 19 to 20, than women who participated. However, the most frequent age range of both males and females was ages 35 to 49. It is also worth noting that the only participants ages 16 to 17 were male. It is at these younger ages that we see more discrepancies between male and female participants. But starting at age 21, the male-female distribution evens out a bit. 


### Problem 4: Substance use

For which of the three substances included in the dataset (marijuana, alcohol, and cigarettes) do individuals tend to use the substance earlier?
```{r}
par("mar")
par(mar=c(1,1,1,1))
par(mfrow=c(3,1))
hist(dat$mjage, main="First use Marijuana", xlab="Age", ylab="Frequency of Age")
hist(dat$cigage, main="First smoked Cigarettes", xlab="Age", ylab="Frequency of Age")
hist(dat$iralcage, main="First tried Alcohol", xlab="Age", ylab="Frequency of Age")
```


Based on the histograms, individuals tend to use alcohol earlier. It had the most frequent values of ages 5 to 10 compared to marijuana and cigarettes. 

### Problem 5: Sexual attraction

What does the distribution of sexual attraction look like? Is this what you expected?
```{r}
#install.packages("magrittr") # package installations are only needed the first time you use it
#install.packages("dplyr")    # alternative installation of the %>%
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)    # alternatively, this also loads %>%
dat$sexatract <- dat$sexatract %>% na_if(., "99")
hist(dat$sexatract, main="Sexual Attraction", xlab="Attraction", ylab="Frequency")
```


After removing the outlying values of 99, which indicates a skip of data, the histogram peaks at a value of 1. This means that the majority of respondents reported that they are only attracted to the opposite sex. Very few respondents answered that they were attracted, or mostly attracted, to the same sex. At least one respondent reported that they were unsure. I would have expected more variation in attraction with higher frequencies of people attracted to the same sex. This could be explained by the fact that the sample size is relatively small, and therefore may not be entirely representative of the U.S. population in terms of sexual attraction. 

What is the distribution of sexual attraction by gender? 
```{r}
tab.sexatract <- table(dat$irsex, dat$sexatract)
barplot(tab.sexatract,
        main = "Stacked barchart",
        xlab = "Attraction category", ylab = "Frequency",
        legend.text = rownames(tab.sexatract),
        beside = FALSE) # Stacked bars (default)
```


According to the barchart more males than females reported that they were only attracted to the opposite sex. Additionally, the majority of males and females reported that they were only attracted to the opposite sex. More females than males reported that they were mostly attracted to the opposite sex or were equally attracted to both sexes. More males than females reported that they were mostly or only attracted to members of the same sex. Additionally, the respondent who reported they were unsure was male. 


### Problem 6: English speaking

What does the distribution of English speaking look like in the sample? Is this what you might expect for a random sample of the US population?
```{r}
hist(dat$speakengl, main="Speak English", xlab="How well", ylab="Frequency")
```


Most of the respondents, over 150, reported that they speak English very well, with few simply answering well, and even fewer reporting that they do not speak it well. In a random sample of the U.S. population, I would expect that more people would answer that they do not speak English very well, creating more variation in the histogram. Due to the large number of immigrants from all over the world it would make sense for more people within the whole population to not feel as confident in their English speaking. 

Are there more English speaker females or males?
```{r}
tab.sexengl <- table(dat$irsex, dat$speakengl)
barplot(tab.sexengl,
        main = "Stacked barchart",
        xlab = "English category", ylab = "Frequency",
        legend.text = rownames(tab.sexengl),
        beside = FALSE) # Stacked bars (default)
```


According to the barchart, those who speak English very well are almost evenly distributed between males and females. However, more males than females reported that they speak English well, indicating that overall, more males than females speak English within the sample. Additionally, more females than males reported that they do not speak English well. 





# Exam 1
## Elizabeth Stoner
## 10/04/2021
output: html_document


### Instructions

a. Create a folder in your computer (a good place would be under Crim 250, Exams). 

b. Download the dataset from the Canvas website (fatal-police-shootings-data.csv) onto that folder, and save your Exam 1.Rmd file in the same folder.

c. Download the README.md file. This is the codebook. 

d. Load the data into an R data frame.
```{r}
library(readr)
fatal_police_shootings_data <- dat<-read_csv("fatal-police-shootings-data.csv")
```


### Problem 1 (10 points)

a. Describe the dataset. This is the source: https://github.com/washingtonpost/data-police-shootings . Write two sentences (max.) about this.

__The dataset describes information about fatal police shootings in the US from 2015 to 2021. There are 6594 subjects included in teh dataset.__

b. How many observations are there in the data frame?
```{r}
names(dat) #Lists the names of all the variables in dat
```

__There are 17 categories used to observe 6594 individuals, so 6594 observations.__

c. Look at the names of the variables in the data frame. Describe what "body_camera", "flee", and "armed" represent, according to the codebook. Again, only write one sentence (max) per variable.
```{r}

```

__"Body camera" means whether news reports indicated that the officer was wearing a body camera at the time of the incident, which may have recorded some of the events. "Flee" means whether news reports said that the subject was moving away from the officer. "Armed" indicates whether the subject was armed with some sort of object that the officer believed could have caused harm.__

d. What are three weapons that you are surprised to find in the "armed" variable? Make a table of the values in "armed" to see the options.
```{r}
table(dat$armed)
```

__I find it most surprising to see an air conditioner, wasp spray, and a microphone in the "armed" variable.__

### Problem 2 (10 points)

a. Describe the age distribution of the sample. Is this what you would expect to see?
```{r}
hist(dat$age, main="Age", xlab="Subject Age", ylab="Frequency")
```

__There seems to be a higher frequency of ages 20 to 40, so the plot has a right skew. It makes sense that people from these age ranges may be more likely to be involved in police incidents. They are able to be more active, unlike many in their eighties. There is a rather steady decline starting from age 40 going to 80 in the ages of those involved in fatal police shootings. This is also what I would expect. The older people get, the less likely they are to engage in criminal activities, or activities that could be mistaken for criminal activities. Older people may also seem less threatening, so officers may be less inclined to use deadly force on them.__

b. To understand the center of the age distribution, would you use a mean or a median, and why? Find the one you picked.
```{r}
summary(dat$age)
```

__To find the center of a distribution it is best to use the median, which is the middle value of a dataset. For this age distribution the median is 35.__

c. Describe the gender distribution of the sample. Do you find this surprising?
```{r}
table(dat$gender)
barplot(table(dat$gender), main="Gender") #Organizing the data into a barplot to visualize the distribution
```

__There are 6298 males and 293 females in the dataset. While I would expect there to be more males than females, I am very surprised that there are over twenty times more males than females involved in fatal police shootings.__


### Problem 3 (10 points)

a. How many police officers had a body camera, according to news reports? What proportion is this of all the incidents in the data? Are you surprised that it is so high or low?

```{r}
table(dat$body_camera)
```

__According to news reports only 910 officers had a body camera to capture parts of the incidents. This is less than one sixth of all of the incidents in the data. I am quite surprised at how low it is. I would also expect that in recent years there would be a greater increase in body camera use; however, looking at the dataset that does not seem to be the case.__

b. In  how many of the incidents was the victim fleeing? What proportion is this of the total number of incidents in the data? Is this what you would expect?
```{r}
table(dat$flee)
```

__The victim fled via car in 1058 incidents and on foot in 845 incidents.So, there were 1903 cases in which the victim fled. There is also an other category applied to 248 cases. However, this category could simply be missing values or other occurrences that are difficult to classify as either fleeing or not fleeing. In 3925 cases the victim was not fleeing. This is about 60% of the cases. I would expect more victims to flee so that there was a more even distribution. However, I suppose it might make sense for officers to interpret victims standing their ground to be more threatening than those trying to run away.__



### Problem 4 (10 points) -  Answer only one of these (a or b).

a. Describe the relationship between the variables "body camera" and "flee" using a stacked barplot. What can you conclude from this relationship? 

*Hint 1: The categories along the x-axis are the options for "flee", each bar contains information about whether the police officer had a body camera (vertically), and the height along the y-axis shows the frequency of that category).*

*Hint 2: Also, if you are unsure about the syntax for barplot, run ?barplot in R and see some examples at the bottom of the documentation. This is usually a good way to look up the syntax of R code. You can also Google it.*


```{r}
tab.bodycameraflee<-table(dat$body_camera, dat$flee) #Making a table for both variables
barplot(tab.bodycameraflee, main="Body Camera-Flee Relationship", xlab="Flee", ylab="Body Camera Frequency", legend.text=rownames(tab.bodycameraflee),
        beside = FALSE) #Stacked barplot code using the tables
```

__From looking at the barplot it is clear that in the majority of cases the officers were not reported to have body cameras. For the cases in which the victims fled, very few officers were reported to have body cameras. The officers were reported to have body cameras mostly when the victims were not fleeing. As this is the largest category, it makes sense that it would also happen to have the most instances of officers having body cameras. This barplot might suggest that when the victims fled, officers perhaps turned off or removed their body cameras. Or the footage could be obstructed due to the action. It makes sense that it is easiest to capture footage of a victim who is stationary. However, the data does not explain the staggering lack of body camera use in all of the categories. Even in the not fleeing category very few officers were reported to have body cameras. It should also be noted that the Other category is likely missing data about whether the victim fled or not. __

b. Describe the relationship between age and race by using a boxplot. What can you conclude from this relationship? 

*Hint 1: The categories along the x-axis are the race categories and the height along the y-axis is age.* 

*Hint 2: Also, if you are unsure about the syntax for boxplot, run ?boxplot in R and see some examples at the bottom of the documentation. This is usually a good way to look up the syntax of R code. You can also Google it.*


```{r}

```

__Your answer here.__






### Extra credit (10 points)

a. What does this code tell us? 

```{r, eval=FALSE}
mydates <- as.Date(dat$date) #renames the date category as "mydates"
head(mydates) #finds the first part of the vector mydates
(mydates[length(mydates)] - mydates[1]) #sets the length of the vector list for mydates
```
__This code tells us about the values in the vector list renamed to "mydates".__

b. On Friday, a new report was published that was described as follows by The Guardian: "More than half of US police killings are mislabelled or not reported, study finds." Without reading this article now (due to limited time), why do you think police killings might be mislabelled or underreported?
__It would be easy to mislabel police killings due to the lack of body cameras worn by officers, which could provide helpful information to understand the facts of teh case. Instead, it is likely a "he-said-she-said" situation in which bystanders and families of the victims feel that their word will not be taken seriously.__

c. Regarding missing values in problem 4, do you see any? If so, do you think that's all that's missing from the data?
__There are around 248 missing values listed as "other" in the flee category. However, it is likely that there is more that could be missing from the data. For example, looking at the dataset shows missing data in the armed category__


# Assignment 3
## Elizabeth Stoner
## 10/23/2021
output:
  pdf_document: default
  html_document: default


**Collaborators: Carmen Avery and Halle Wasser**.

This assignment is due on Canvas on Wednesday 10/27/2021 before class, at 10:15 am. Include the name of anyone with whom you collaborated at the top of the assignment.

Submit your responses as either an HTML file or a PDF file on Canvas. Also, please upload it to your website.

Save the file (found on Canvas) crime_simple.txt to the same folder as this file (your Rmd file for Assignment 3).

Load the data.
```{r}
library(readr)
library(knitr)
dat.crime <- read_delim("crime_simple.txt", delim="\t")
```

This is a dataset from a textbook by Brian S. Everitt about crime in the US in 1960. The data originate from the Uniform Crime Report of the FBI and other government sources. The data for 47 states of the USA are given. 

Here is the codebook:

R: Crime rate: # of offenses reported to police per million population

Age: The number of males of age 14-24 per 1000 population

S: Indicator variable for Southern states (0 = No, 1 = Yes)

Ed: Mean of years of schooling x 10 for persons of age 25 or older

Ex0: 1960 per capita expenditure on police by state and local government

Ex1: 1959 per capita expenditure on police by state and local government

LF: Labor force participation rate per 1000 civilian urban males age 14-24

M: The number of males per 1000 females

N: State population size in hundred thousands

NW: The number of non-whites per 1000 population

U1: Unemployment rate of urban males per 1000 of age 14-24

U2: Unemployment rate of urban males per 1000 of age 35-39

W: Median value of transferable goods and assets or family income in tens of $

X: The number of families per 1000 earning below 1/2 the median income


We are interested in checking whether the reported crime rate (# of offenses reported to police per million population) and the average education (mean number of years of schooling for persons of age 25 or older) are related. 


### 1. How many observations are there in the dataset? To what does each observation correspond?

```{r}
summary(dat.crime)
```

__There are 47 observations in the dataset, which correspond to 47 of the USA states and their 1960 crime data. There are 14 columns of variables providing information for each observation.__

### 2. Draw a scatterplot of the two variables. Calculate the correlation between the two variables. Can you come up with an explanation for this relationship?

```{r, fig.width=6, fig.height=4}
plot(dat.crime$Ed, dat.crime$R, main="Relationship between Average Education and Crime Rate for 47 States", xlab="Average Education", ylab="Crime Rate", cex.main=1)
```



```{r}
cor(dat.crime$Ed, dat.crime$R)
```

__The correlation between the two variables is 0.3228349, which is a positive correlation. According to the codebook, R, the crime rate, is the number of offenses reported to police per million population, and Ed is the average years of schooling multiplied by ten for people of 25 years or older. So, as the average amount of schooling increases, so does the crime rate. A possible explanation for this relationship is that areas where people have higher levels of education may be better targets for those committing crimes. Education level is often correlated with better jobs and financial success, so perhaps areas with these people may be home to more criminal activity. Also, because R is specifically number of crimes reported, then maybe people with higher education levels are also more likely to report crimes to the police; however, this hypothesis cannot be proven by this data alone.__

### 3. Regress reported crime rate (y) on average education (x) and call this linear model `crime.lm` and write the summary of the regression by using this code, which makes it look a little nicer `{r, eval=FALSE} kable(summary(crime.lm)$coef, digits = 2)`.

```{r} 
#install.packages("kableExtra")
crime.lm <- lm(formula=R~Ed, data=dat.crime)
# Remember to remove eval=FALSE above!
summary(crime.lm)
```


### 4. Are the four assumptions of linear regression satisfied? To answer this, draw the relevant plots. (Write a maximum of one sentence per assumption.)

```{r} 
plot(dat.crime$Ed, crime.lm$residuals, main="Residuals vs. x", xlab="x, Education", ylab="Residuals") #Visualizing the data
```


```{r}
plot(crime.lm, which=1) #Plot for the linearity assumption and independence assumption
```


__Based on the plot, the linearity assumption does not look to be met as the red line appears to have a varying pattern that is non-linear. The same can be said for the independence assumption; there is some sort of pattern, so the independence assumption is not met as well as we would like it to be.__

```{r}
plot(crime.lm, which=3) #Plot to test for the equal variance assumption 
```


__Using the scale-location plot, it appears as though the data does not perfectly satisfy the equal variance assumption as the line curves upward then decreases around the x-value of 100, and the distribution of the data points increase around the middle but is thinner near the beginning, suggesting a lack of homoscedasticity.__

```{r}
plot(crime.lm, which=2) #Plot to test the normal population assumption
```


__Based on the normal qq plot there are issues with the normal population assumption, as the plot is light tailed on both ends, meaning that those data points could be outliers, which no longer follow the slope of the line.__

### 5. Is the relationship between reported crime and average education statistically significant? Report the estimated coefficient of the slope, the standard error, and the p-value. What does it mean for the relationship to be statistically significant?

__The estimated coefficient of the slope is 1.1161, which indicates the effect that education has on the reported crime rate. The standard error is 0.4878, which indicates the potential variance of crimes actually reported. The p-value is 0.0269 for the slope, which indicates that there is only a weak relationship between reported crime and average education. This means that the relationship is not statistically significant. A statistically significant relationship would be such that the relationship is unlikely to occur randomly, or unlikely to occur given the null hypothesis. Therefore, this relationship is more likely to have occurred due to randomness or some other factor.__

### 6. How are reported crime and average education related? In other words, for every unit increase in average education, how does reported crime rate change (per million) per state?

__Based on the data, for every increase in one unit of average education, number of offenses reported to the police increases by 1.1161 per million per state.__

### 7. Can you conclude that if individuals were to receive more education, then crime will be reported more often? Why or why not?

__Using the data provided we cannot conclude that if individuals were to receive more education, then crime would be reported more often. The statistical relationship is weak, at best. And, even so, that relationship does not automatically indicate causality. There could be other factors that have yet to be measured that could influence the perceived relationship. For example, community relations with local law enforcement is an important factor in a willingness to report crimes, but is not observably related to education levels. So, in order to infer any further relationship, more information and more data are needed.__

